Prelim report

### 2021-02-10 docs/ms/ms.RMD
preliminary documents and first draft of figures.

# Introduction
<!--- warming and effects in food webs --->
Temperature is an important abiotic variable with increasing relevance as increasing global temperatures are altering the diversity, structure, and functioning of Earth's ecosystems [@brose2012][@walther2002][]. Temperature'smak effect on ecosystem manifest through complex of direct and indirect pathways such as, shifting species ranges [][] and subsequent changes to local and regional communities [], species adaptations [@gibert2017], and through effects on individual metabolic rates [@brown2004][@arrhenius18].While a growing body of theoretical and empirical study has enhanced our knowledge of temperature-mediated changes to ecosystems [][][], general patterns are uncertain and empirical studies often idiosyncratic, especially at higher levels of organizations such as communities and food webs [@walter2012][@woodward2010]. This uncertainty arises because the influence of environmental temperature on ecosystems occurs through complex direct and indirect pathways, s Identifying general effects of temperature is further complicated by the fact that natural variation in temperature often exists concurrently with other environmental gradients (e.g., light, water, and/or nutrient availability). Understanding the effects of temperature...alterations to ecosystem function and the services they provide. 
<!--- Speeding up of food web: Species levels--->
The provision and maintenance of ecosystem functions are underpinned by the network of species interactions across which energy, materials, and information are exchanged (e.g., food webs; [@deruiter1995][@thompson2012]. Characteristics The magnitude The response of ecosystem function to environmental drivers and variability <!---Ecosystem functionality---> is tied to the magnitudes and relative distributions of energy fluxes among species and trophic levels [@may1972][@barnes2018]. The complexity in responses to warming arise from 
<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
Here, we measured the patterning and distribution of organic matter fluxes within invertebrate food webs across a natural stream temperature gradient (~5 - 28$^\circ$C). Previous research in these streams has shown a strong positive effect of temperature on primary production both among streams [@demars2011][@padfield2017] and within streams seasonally [@gorman2012][@hood2018]. The dynamics of basal resource supply has a strong control on consumer energy demand [@junker2020], as such, we predicted total annual OM fluxes to consumer to scale with among stream patterns in primary production and consumer energy demand, and therefore, increase with temperature across streams. Further, we predicted that food web OM fluxes will 'speed up' with increasing temperature through increases in the importance of relatively faster, higher turnover consumers. Within streams, the speeding up of ecosystems will lead to earlier food web fluxes in warmer compared to colder streams.

>>>>

# Introduction
<!--- warming and effects in food webs --->
Temperature is an important abiotic variable with increasing relevance  increasing global temperatures are altering the diversity, structure, and functioning of Earth's ecosystems [@brose2012][@walther2002][]. Temperature's effect on ecosystems manifest through complex of direct and indirect pathways such as, shifting species ranges [][] and subsequent changes to local and regional communities [], species adaptations [@gibert2017], and through effects on individual metabolic rates [@brown2004][@arrhenius18].While a growing body of theoretical and empirical study has enhanced our knowledge of temperature-mediated changes to ecosystems [][][], general patterns are uncertain and empirical studies often idiosyncratic, especially at higher levels of organizations such as communities and food webs [@walter2012][@woodward2010]. This uncertainty arises because the influence of environmental temperature on ecosystems occurs through complex direct and indirect pathways, which are concurrently influenced by other environmental gradients (e.g., light, water, and/or nutrient availability). Yet, identifying general effects of temperature is a necessary step ...alterations to ecosystem function and the services they provide. 
<!--- Speeding up of food web: Species levels--->
The provision and maintenance of ecosystem functions are underpinned by the network of species interactions across which energy, materials, and information are exchanged (e.g., food webs; [@deruiter1995][@thompson2012]. Characteristics The magnitude The response of ecosystem function to environmental drivers and variability <!---Ecosystem functionality---> is tied to the magnitudes and relative distributions of energy fluxes among species and trophic levels [@may1972][@barnes2018]. The complexity in responses to warming arise from 
<!--- Connect traits to the 'speed' of ecosystem energy fluxes--->
Here, we measured the patterning and distribution of organic matter fluxes within invertebrate food webs across a natural stream temperature gradient (~5 - 28$^\circ$C). Previous research in these streams has shown a strong positive effect of temperature on primary production both among streams [@demars2011][@padfield2017] and within streams seasonally [@gorman2012][@hood2018]. The dynamics of basal resource supply has a strong control on consumer energy demand [@junker2020], as such, we predicted total annual OM fluxes to consumer to scale with among stream patterns in primary production and consumer energy demand, and therefore, increase with temperature across streams. Further, we predicted that food web OM fluxes will 'speed up' with increasing temperature through increases in the importance of relatively faster, higher turnover consumers. Within streams, the speeding up of ecosystems will lead to earlier food web fluxes in warmer compared to colder streams.

### 2021-02-10 docs/prelim-doc.Rmd
preliminary documents and first draft of figures.

---
title: "Prelim report"
author: "Junker"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document:
      keep_md: true
      toc: true
      toc_float: true
      toc_depth: 3
      numbered_sections: true
---

@@

source("./packages.R")
stream_order = factor(c("hver", "st6","st9", "st7","oh2","st14"))#stream ordering
stream_temp_labels = c("27.2","17.6","11.2","5.8","5.5","5.0")#stream annual temperature labels
oce_temp_pos = c(256,212,168,124,80,1)#color positions in 'temperature' list of ocecolors

drake::loadd("spp_rankings")

@@

```{r pb historgram, echo = FALSE, warnings = FALSE}
pb_df = spp_rankings[['PB_spp_rank']] %>% bind_rows %>% dplyr::filter(pb_y_mean >0) %>% dplyr::mutate(site_id = factor(site_id, levels = rev(stream_order)))
bio_df = spp_rankings[['bio_spp_rank']] %>% bind_rows %>% dplyr::filter(bio_mg_m_mean >0 | is.na(bio_mg_m_mean)) %>% dplyr::mutate(site_id = factor(site_id, levels = rev(stream_order)))
M_df = spp_rankings[['M_spp_rank']] %>% bind_rows %>% dplyr::filter(M_mg_ind_mean >0 | is.na(M_mg_ind_mean)) %>% dplyr::mutate(site_id = factor(site_id, levels = rev(stream_order)))

pb_boxplot = ggplot(pb_df,aes(x = site_id, y = log10(pb_y_mean))) + geom_violin(aes(fill = site_id), alpha = 0.5) +
  geom_jitter(aes( fill = site_id),color = 'black', size = 2, shape = 21, alpha = 0.5) +
  scale_fill_manual(values = rev(ocecolors[['temperature']][oce_temp_pos]), labels = rev(stream_temp_labels)) +
  theme(axis.title.x = element_blank(), legend.title = element_blank(), legend.position = c(0,0), legend.justification = c(0,0),
        legend.background = element_rect(fill = 'transparent', colour = NA))
# pb_boxplot

M_boxplot = ggplot(M_df,aes(x = site_id, y = log10(M_mg_ind_mean))) + geom_violin(aes(fill = site_id), alpha = 0.5) +
  geom_jitter(aes( fill = site_id),color = 'black', size = 2, shape = 21, alpha = 0.5) +
  scale_fill_manual(values = rev(ocecolors[['temperature']][oce_temp_pos]), labels = rev(stream_temp_labels)) +
  theme(axis.title.x = element_blank(), legend.title = element_blank(), legend.position = "none")

bio_boxplot = ggplot(bio_df,aes(x = site_id, y = log10(bio_mg_m_mean))) + geom_violin(aes(fill = site_id), alpha = 0.5) +
  geom_jitter(aes( fill = site_id),color = 'black', size = 2, shape = 21, alpha = 0.5) +
  scale_fill_manual(values = rev(ocecolors[['temperature']][oce_temp_pos]), labels = rev(stream_temp_labels)) +
  theme(axis.title.x = element_blank(), legend.title = element_blank(),legend.position = "none")

grid.arrange(pb_boxplot, M_boxplot, bio_boxplot, ncol = 3)

@@

@@

<details><summary>Reproducibility receipt</summary>

@@

>>>>

---
title: "Prelim report"
author: "Junker"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    keep_md: yes
    toc: yes
    toc_float: yes
    numbered_sections: yes
bibliography: refs.bib
---

@@

source("./packages.R")
stream_order = factor(c("hver", "st6","st9", "st7","oh2","st14"))#stream ordering
stream_temp_labels = c("27.2","17.6","11.2","5.8","5.5","5.0")#stream annual temperature labels
oce_temp_pos = c(256,212,168,124,80,1)#color positions in 'temperature' list of ocecolors
library(knitr)

@@

First, thing I did was find a way to model and propagate error of diets across sites. To do so, I build a bayesian random effects model that incorporated varying seasons, taxon, and streams and modeled the multinomial diet response using a dirichlet family distributional model. The model priors assume equal contribution of all diet items, which is conservative, but it outputs posterior distributions of diet proportions. These we can draw from for each taxa and each bootstrap to incorporate uncertainty through the flux estimates. I only estimated this for non-predators and then predators assumed to have a diet of 100% animal material. The posterior distributions look something like the plot below. This is subsampled to allow for a more manageable plot file size, but it captures the distributions.

### Stream-level posterior distributions of diet contributions

```{r diet plots, echo = FALSE, warning=FALSE}

drake::loadd(prelim_diet_figures)

prelim_diet_figures

@@

When we look at the distribution of OM flux to taxa, we see varying evenness in the distributions of OM fluxes. One way to view this is rank-production plots:

```{r rank-production plots, echo = FALSE, warning=FALSE, height = 8, width = 8, units = 'in'}
drake::loadd(annual_spp_flux_fig)

annual_spp_flux_fig

This gives the impression that the evenness of production is highest in the mid range temperatures. However, evenness is not as straightforward because evenness is related to species richness in almost all formulations of it. A consensus seems to be emerging that evenness must be corrected for richness to compare across sites [@chao2019]. Once we correct for this we see evenness is not lowest at higher temperatures, but in the middle when we account for species richness. This is based on the Gini index which measures inequality and when standardized ranges from 0 to 1.

```{r gini table, echo = FALSE, warning=FALSE}

drake::loadd(gini_analysis)
drake::loadd(environment_data)

ann_temps = environment_data[["ann_stream_temps"]]

gini_summ = gini_analysis[["stream_gini_summary"]] %>% dplyr::select(c(site, contains("mean"), contains("quant2.5"), contains("quant97.5"))) %>% left_join(ann_temps %>% dplyr::select(site = "SITE", tempC), by = "site") %>% dplyr::mutate(tempc = round(tempC, 1)) %>% dplyr::select(site, tempC, contains('Non-normalized'), contains("Normalized")) %>% dplyr::arrange(desc(tempC))
knitr::kable(gini_summ, digits = 2)
```

Another way to view this is to look at the accumulation of flux across the accumulation of species standardize 0-1. This is often viewed as a Lorenz curve. Here we order species by annual flux, $x_i$ is the annual flux of species *i*, such that $x_1 < x_2 < ... < x_n$. If every species contributes an equal amount to production the Lorenz curve falls along the solid 1:1 line on the plot. The bold colored lines represent the mean value and each grey line is an individual bootstrapped estimate.

### Lorenz curves of cumulative flux with species

```{r lorenz of flux, echo = FALSE, warning=FALSE}
drake::loadd(lorenz_trait_fig)

lorenz_trait_fig[[1]]

Here we see that total community flux accumulates differently across the streams. There are a few instances when it looks like abrupt 'jumps' in cumulative flux. These are mostly driven by species that dominate biomass. When we look at how flux accumulates with biomass we see the curves fall near the 1:1 lines, which is expected since standing biomass strongly influences production. In st14 we see lots of variation driven by empty samples and low sample size in the winter.

### Lorenz curve of cumulative flux v. biomass

```{r lorenz biomass, echo = FALSE, warning=FALSE}

lorenz_trait_fig[[4]]

This allows us an opportunity to look how total community flux accumulates across different species traits in a standardized way. So rather than the distribution differences attributed to different absolute magnitudes they are attributed to the relative magnitudes. So how does temperature modify how flux accumulates with species traits such as annual P:B and mean body size. First, a look at the distribution of traits across taxa in the streams.

```{r pb histogram, echo = FALSE, warnings = FALSE}

drake::loadd(spp_trait_histograms)

grid.draw(spp_trait_histograms)

So how does flux accumulate with annual *P:B* across sites? Here, we order each species by *P:B*, such that $pb_1 < pb_2 < ... < pb_n$. We see the most striking differences at the warmest and coldest sites. At hver, OM flux is disproportionately skewed towards taxa with higher annual *P:B* and in st14 flux is disproportionately skewed towards low *P:B* taxa.

### Accumulation of flux with P:B across sites

```{r pb lorenz, echo = FALSE, warning=FALSE}
lorenz_trait_fig[[2]]
```

So how does flux accumulate with mean body size, *M*, across sites? Here, we order each species by *M*, such that $M_1 < M_2 < ... < M_n$. This give a complementary story to *P:B*, flux in hver is disproportionately skewed towards lower *M*, and st14 skewed towards larger relative *M*. We are also able to glean a bit more information from some of the mid temperature streams. In st9 (11.2 $^\circ$C) the largest taxa also has the largest flux (and biomass)

### Accumulation of flux with M across sites

```{r M lorenz, echo = FALSE, warning=FALSE}
lorenz_trait_fig[[3]]
```

@@

<details>

<summary>

Reproducibility receipt

</summary>

@@

## References
